---
title: "Poster Data Analysis"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

# Description
This notebook presents the analysis of whole genome data for the poster "Quinolone resistance mechanisms found in E. coli from four animal species in Norway
", presented at the ASM NGS Conference in Washington D.C. September 2018.

```{r, include = FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr)

```

# Functions
This chunk lists the functions used in this notebook.
```{r}
# Functions

func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

## COnfidence intervals
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

# Creates a correlation matrix from 1/0 data from MIC values
create_corr_matrix <- function(df) {
  df <- df %>%
    as.matrix %>%
    cor %>%
    as.data.frame %>%
    rownames_to_column(var = "var1") %>%
    gather(var2, value, -var1) %>%
    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
           var2 = gsub("(.*?)_R$", "\\1", var2))
  return(df)
}

# Creates a data frame with p-values for comparing presence of specific genes between animal species
do_chisq_test <- function(df) {
  df <- df %>%
    slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
          filter(Var2 < Var1) %>% t) %>%
    mutate(test = rep(1:(n()/2), each = 2)) %>%
    group_by(test) %>%
    do(data_frame(gene = .$key,
                  test = first(.$test),
                  species1 = first(.$species),
                  species2 = last(.$species),
                  data = list(matrix(c(.$Absent,
                                       .$Present),
                                     ncol = 2)))) %>%
    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
    mutate(p.value = map_dbl(chi_test, "p.value"),
           dupl = duplicated(test)) %>%
    filter(dupl == FALSE) %>%
    select(gene, species1, species2, p.value) %>%
    filter(p.value < 0.05)
  return(df)
}
```

# Import data
This chunk imports data from MLST and ARIBA AMR gene detection results. Isolate data is also imported.
```{r}
# Isolate data, including DNA extraction results, species,
# method of isolation, and sequencing methodology
isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

# Total MIC profile for all isolates in this study
mic_data <- read.table("../data/mic_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

# ARIBA MLST results for each isolate
mlst_results <- read.table("../data/mlst_results.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

# ARIBA AMR gene results (megares database, for mutations in intrinsic genes)
megares_results <- read.table("../data/megares_results.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

# ARIBA AMR gene results (resfinder database, for acquired genes)
resfinder_results <- read.table("../data/resfinder_results.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

species_data <- select(isolate_data, c(id, species))

total_data <- isolate_data %>%
  left_join(mic_data, by = "id") %>%
  left_join(mlst_results, by = "id") %>%
  rename("m_gyrB" = gyrB) %>%
  left_join(resfinder_results, by = "id") %>%
  select(-type) %>%
  left_join(megares_results, by = "id") %>%
  select(-type)
```

# Tables

## Samples per species
```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  count()
per_species
```

## Samples per sequencing method
```{r}
per_seq <- isolate_data %>%
  group_by(sequencer, libprep, read_length) %>%
  count()
per_seq
```

## MLST Data

### Total MLST 
```{r}
total_mlst <- mlst_results %>%
  left_join(species_data, by = "id") %>%
  group_by(ST) %>%
  count()
total_mlst
```

### MLST per animal species
```{r}
mlst_per_species <- mlst_results %>%
  left_join(species_data, by = "id") %>%
  group_by(ST, species) %>%
  count()
mlst_per_species
```



# Figures
## Percent presence of acquired genes in total
```{r}
resfinder_results %>%
  gather(gene, result, -c(id, type)) %>%
  group_by(gene, result) %>%
  count() %>%
  ungroup() %>%
  mutate(result = if_else(result == 1, "Present", "Absent")) %>%
  spread(result, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1)) %>%
  ggplot(aes(gene, Percent))+
  geom_col(color = "black")+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6)+
  labs(y = "Percent (%) of isolates")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())
```

## Percent presence of mutations in intrinsic genes in total
```{r}
megares_results %>%
  gather(gene, result, -c(id, type)) %>%
  group_by(gene, result) %>%
  count() %>%
  ungroup() %>%
  mutate(compl = complete.cases(.),
         result = if_else(result == 1, "Present", "Absent")) %>%
  filter(compl == TRUE) %>%
  select(-compl) %>%
  spread(result, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1)) %>%
  ggplot(aes(gene, Percent))+
  geom_col(color = "black")+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6)+
  labs(y = "Percent (%) of isolates")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())
```

## Acquired genes per animal species
### Total acquired genes per animal species
```{r}
gene_names <- names(resfinder_results)

gene_names <- gene_names[-c(1,2)]

resfinder_results %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, gene_names) %>%
  group_by(key, value, species) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1)) %>%
  ggplot(aes(key, Percent, fill = species))+
  geom_col(color = "black",
           position = position_dodge(1))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6,
                position = position_dodge(1))+
  scale_fill_brewer(palette = "Paired")+
  labs(y = "Percent (%) of isolates")+
  guides(fill = FALSE)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())+
  facet_wrap(~species)
```

### Qnr-genes per animal species
```{r}
qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")

qnr_data <- resfinder_results %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, qnr_genes) %>%
  group_by(key, value, species) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Present + Absent,
         Percent = round(Present/Total*100, 1),
         lwr = round(get_binCI(Present, Total)[1], 1),
         upr = round(get_binCI(Present, Total)[2], 1))

ggplot(qnr_data, aes(key, Percent, fill = species))+
  geom_col(color = "black",
           position = position_dodge(0.9))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6,
                position = position_dodge(0.9))+
  scale_fill_brewer(palette = "Paired")+
  labs(y = "Percent (%) of isolates",
       fill = NULL)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())
```



## Mutations in intrinsic genes per animal species
```{r}
mut_names <- names(megares_results)

mut_names <- mut_names[-c(1,2)]

megares_results %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, mut_names) %>%
  group_by(key, value, species) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Mutant", "Wild Type"),
         compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  spread(value, n, fill = 0) %>%
  rowwise() %>%
  mutate(Total = Mutant + `Wild Type`,
         Percent = round(Mutant/Total*100, 1),
         lwr = round(get_binCI(Mutant, Total)[1], 1),
         upr = round(get_binCI(Mutant, Total)[2], 1)) %>%
  ggplot(aes(key, Percent, fill = species))+
  geom_col(color = "black",
           position = position_dodge(1))+
  geom_errorbar(aes(ymin = lwr, ymax = upr),
                width = 0.6,
                position = position_dodge(1))+
  scale_fill_brewer(palette = "Paired")+
  labs(y = "Percent (%) of isolates")+
  guides(fill = FALSE)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, size = 8, hjust = 1, vjust = 0.4),
        axis.title.x = element_blank())+
  facet_wrap(~species)
```

## Correlation plot for mutations in intrinsic genes
```{r}
corr_megares <- megares_results %>%
  select(-c(type, acrB)) %>%
  left_join(species_data, by = "id") %>%
  mutate(compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  split(., f = .$species) %>%
  lapply(., function(x) select(x, -c(species, id, compl)))

mut_corr_matrices <- lapply(corr_megares, function(x) create_corr_matrix(x)) %>%
  bind_rows(.id = "species")

cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")

ggplot(mut_corr_matrices, aes(var1, var2, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradientn(colors=cols,
                       na.value="grey95",
                       limits=c(-1, 1),
                       guide=guide_colourbar(ticks=T,nbin=10,
                                             barheight=.5,label=T, 
                                             barwidth=10))+
  labs(fill = NULL)+
  theme_classic()+
  theme(axis.text = element_text(size = 7),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
        axis.title = element_blank(),
        legend.position="bottom",
        legend.justification="center",
        legend.direction="horizontal")+
  coord_fixed()+
  facet_wrap(~species)
```

## Correlation plot for acquired genes
```{r message=FALSE, warning=FALSE}
corr_resfinder <- resfinder_results %>%
  select(-type) %>%
  left_join(species_data, by = "id") %>%
  split(., f = .$species) %>%
  lapply(., function(x) select(x, -c(species, id)))

acquired_corr_matrices <- lapply(corr_resfinder, function(x) create_corr_matrix(x)) %>%
  bind_rows(.id = "species")

cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")

ggplot(acquired_corr_matrices, aes(var1, var2, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradientn(colors=cols,
                       na.value="grey95",
                       limits=c(-1, 1),
                       guide=guide_colourbar(ticks=T,nbin=10,
                                             barheight=.5,label=T, 
                                             barwidth=10))+
  labs(fill = NULL)+
  theme_classic()+
  theme(axis.text = element_text(size = 6),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
        axis.title = element_blank(),
        legend.position="bottom",
        legend.justification="center",
        legend.direction="horizontal")+
  coord_fixed()+
  facet_wrap(~species)
```

## MIC plots

### MIC Distribution for generated groups
```{r message=FALSE, warning=FALSE}
selected_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")
selected_mut <- c("gyrA","gyrB","parC","parE","marR")

micplot_data <- mic_data %>%
  select(id, CIP) %>%
  left_join(megares_results, by = "id") %>%
  left_join(resfinder_results, by = "id") %>%
  select(-c(acrB, type.x, type.y)) %>%
  mutate(compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  select_at(vars(id, CIP,selected_mut, selected_genes)) %>%
  mutate(none = if_else(rowSums(.[,3:ncol(.)]) != 0, 0, 1)) %>%
  group_by_at(vars(selected_mut, selected_genes, none)) %>%
  { mutate(ungroup(.), group = group_indices(.)) } %>%
  gather(gene, value, -c(id, CIP, group)) %>%
  mutate(gene = ifelse(value == 0 & gene != "none", NA, gene),
         value = ifelse(gene == "none" & value == 0, NA, value)) %>%
  na.omit()


factor_breaks <- c(0.015, 0.03, 0.06, 0.12, 0.25, 0.5, 1, 2, 4, 8, 16)
  
factor_levels <- c("0.015", "0.03", "0.06", "0.12", "0.25", "0.5", "1", "2", "4", 
                   "8", "16")


p0 <- micplot_data %>%
  group_by(id, group) %>%
  count() %>%
  group_by(group) %>%
  count() %>%
  ggplot(aes(factor(group), nn, fill = factor(group)))+
  geom_col(color = "black")+
  guides(fill = FALSE)+
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)+
  theme_light()+
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

p1 <- micplot_data %>%
  group_by(CIP, group, gene) %>%
  count() %>%
  ggplot(aes(factor(group), CIP, fill = factor(group), size = n))+
  geom_point(pch = 21)+
  scale_y_continuous(trans = "log10",
                     labels = factor_levels,
                     breaks = factor_breaks)+
  guides(fill = FALSE,
         size = FALSE)+
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)+
  theme_light()+
  theme(axis.title = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

p2 <- ggplot(micplot_data,
       aes(factor(group),
           factor(gene),
           fill = factor(group)))+
  geom_point(shape = 23,
             pch = 21,
             size = 2)+
  geom_hline(yintercept = 5.5)+
  guides(fill = FALSE)+
  scale_fill_viridis(discrete = TRUE,
                     direction = -1)+
  labs(y = "Genes")+
  theme_light()+
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 8),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank())

plot_grid(p0, p1, p2, align = "v", nrow = 3, rel_heights = c(1/4, 2/4, 1/4))
```

### MIC correlation plots per animal species

```{r}
test <- mic_data %>%
  select(-antres) %>%
  left_join(species_data, by = "id") %>%
  mutate(SMX_R = ifelse(SMX > 64, 1, 0),
         TMP_R = ifelse(TMP > 2, 1, 0),
         CIP_R = ifelse(CIP > 0.06, 1, 0),
         TET_R = ifelse(TET > 8, 1, 0),
         NAL_R = ifelse(NAL > 16, 1, 0),
         CTX_R = ifelse(CTX > 0.25, 1, 0),
         CHL_R = ifelse(CHL > 16, 1, 0),
         AMP_R = ifelse(AMP > 8, 1, 0),
         GEN_R = ifelse(GEN > 2, 1, 0)) %>%
  select(contains("_R"), species) %>%
  split(., f = .$species) %>%
  lapply(., function(x) select(x, -species))

corr_matrices <- lapply(test, function(x) create_corr_matrix(x)) %>%
  bind_rows(.id = "species")

cols <- c("#e7f0fa","#c9e2f6","#95cbee","#0099dc","#4ab04a",
          "#ffd73e","#eec73a","#e29421","#f05336","#ce472e")

ggplot(corr_matrices, aes(var1, var2, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradientn(colors=cols,
                       na.value="grey95",
                       limits=c(-1, 1),
                       guide=guide_colourbar(ticks=T,nbin=10,
                                             barheight=.5,label=T, 
                                             barwidth=10))+
  labs(fill = NULL)+
  theme_classic()+
  theme(axis.text = element_text(size = 7),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
        axis.title = element_blank(),
        legend.position="bottom",
        legend.justification="center",
        legend.direction="horizontal")+
  coord_fixed()+
  facet_wrap(~species)
```

## Heatmap plot

```{r fig.height=30, fig.width=10, message=FALSE, warning=FALSE}
megares_long <- megares_results %>%
  gather(gene, value, -c(id, type))

resfinder_long <- resfinder_results %>%
  gather(gene, value, -c(id, type))

megares_long %>%
  rbind(resfinder_long) %>%
  mutate(type2 = ifelse(type == "mut", 1, 2),
         value = factor(ifelse(is.na(value) == TRUE, 2, value))) %>%
  spread(id, value, fill = 0) %>%
  gather(key = "id", value = "result", -c(gene, type, type2)) %>%
  left_join(species_data, by = "id") %>%
  mutate(result = ifelse(result == 2, NA, result),
         fillcol = paste(species, result, sep = "_"),
         fillcol = ifelse(is.na(result) == TRUE, NA, fillcol)) %>%
  ggplot(aes(fct_reorder(gene, type2),
             factor(id,
                    levels = unique(.[order(.$species,.$id),"id",drop = TRUE])),
             fill = fillcol,
             alpha = result)) +
  geom_tile(color = "white")+
  geom_vline(xintercept = 14.5)+
  scale_fill_brewer(palette = "Paired",
                    na.value = "#6b6767")+
  theme_minimal()+
  labs(fill = NULL)+
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.3,
                                   size = 7),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_blank(),
        legend.position = "right")+
  guides(alpha = FALSE)+
  coord_fixed()
  
```

# Statistics

## Significance test for the percent presence of acquired genes between animal species
Only significant values will be presented in the table below (p < 0.05).
```{r message=FALSE, warning=FALSE}
acquired_genes_list <- resfinder_results %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, gene_names) %>%
  group_by(key, value, species) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  select(key, species, Present, Absent) %>%
  split(f = .$key)

p_values_acquired <- lapply(acquired_genes_list, function(x) do_chisq_test(x)) %>%
  bind_rows()

p_values_acquired
```

## Significance test for the percent presence of mutations in intrinsic genes between animal species
Only significant values will be presented in the table below (p < 0.05).

```{r message=FALSE, warning=FALSE}
mut_genes_list <- megares_results %>%
  left_join(isolate_data, by = "id") %>%
  gather(key, value, mut_names) %>%
  filter(key != "acrB") %>%
  group_by(key, value, species) %>%
  count() %>%
  ungroup() %>%
  mutate(value = if_else(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  select(key, species, Present, Absent) %>%
  split(f = .$key)

p_values_mut <-  lapply(mut_genes_list, function(x) do_chisq_test(x)) %>%
  bind_rows()

p_values_mut
```
